SSA based global code motion of identical computations

### 2017-04-14 hoist.tex
goals

This paper describes code-motion of identical computations, a technique that
uses the information computed for \PRE{} to detect identical computations but
has a transformation phase whose goal differs from \PRE{}: it removes identical
computations from different branches of execution.  These identical computations
in different branches of execution are not redundant computations at run-time
and the number of run-time computations is not reduced. It is not a redundancy
elimination pass, and thus it has different cost function and heuristics than
\PRE{} or CSE. It is more similar to global code scheduling
\cite{dragonbook,click1995global} in the sense that it will only move
computations. The goals of code hoisting are:

\begin{itemize}
\item to reduce the code size of the program;
\item to improve function inlining heuristics: functions become cheaper to
  inline when their code size decreases because inliner heuristics in \LLVM{}
  depends on instruction count.
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, instruction schedulers can move heavy
  computations earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting expressions out of branches, code hoisting can effectively reduce the
  amount of code to be speculatively executed and can reduce the critical path;
\item to reduce interference/register pressure: by moving computations closer to
  the definitions of their operands in cases where hoisting would reduce the
  interference. For example if all the use operands of an expression of the form
  a = b op c are last usages (a kill) then hoisting will reduce the interference
  by two but increase by one.
\item for SIMD architectures which execute all branches, code hoisting will
  reduce the total number of instructions to be executed.
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted or sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

The main contributions of this paper are:
\begin{itemize}
\item a new optimistic algorithm to hoist computations from branches,
\item cost models to reduce live-range and reduce spills,
\item a technique to maximize hoisting in an optimistic approach by partitioning
  the list of potential candidates sorted by their DFS visit number,
\item performance evaluation of the implementation in \LLVM{},
\item comparison against other algorithms for code hoisting.
\end{itemize}

>>>>

This paper describes code-motion of identical computations, a technique that
uses the information computed for \PRE{} to detect identical computations but
has a transformation phase whose goal differs from \PRE{}: it removes identical
computations from different branches of execution.  These identical computations
in different branches of execution are not redundant computations at run-time
and the number of run-time computations is not reduced. It is not a redundancy
elimination pass, and thus it has different cost function and heuristics than
\PRE{} or CSE. It is more similar to global code scheduling
\cite{dragonbook,click1995global} in the sense that it will only move
computations. The goals of this code motion are:

\begin{itemize}
\item to reduce the code size of the program;
\item to improve function inlining heuristics: functions become cheaper to
  inline when their code size decreases because inliner heuristics in \LLVM{}
  depends on instruction count.
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, instruction schedulers can move heavy
  computations earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting/sinking expressions out of branches, code hoisting can effectively
  reduce the amount of code to be speculatively executed and can reduce the
  critical path;
\item to reduce interference/register pressure: For example if all the use
  operands of an instruction (I) of the form a = b op c are last usages (a kill)
  then hoisting will reduce the interference by two but increase by one across
  the path I gets hoisted.
\item for SIMD architectures, which execute all branches, it will reduce the
  total number of instructions to be executed.
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted or sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

The main contributions of this paper are:
\begin{itemize}
\item a new optimistic algorithm to move computations out of branches,
\item cost models to reduce live-range and reduce spills,
\item a technique to maximize hoisting in an optimistic approach by partitioning
  the list of potential candidates sorted by their DFS visit number,
\item performance evaluation of the implementation in \LLVM{}.
\end{itemize}
